Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm

### 2019-09-29 paper/virtual_cytoplasm_v2.tex
tex update

\documentclass[journal=jacsat,manuscript=article]{achemso}
% \linespread{1.5}
% \usepackage{amssymb,amsmath,graphics,epsfig}
\usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

Crowding ... ~\cite{feig2017crowding} ~\citep{ostrowska2019modeling}

-> Crowding affects hydration around proteins ~\cite{Harada2012a,king2013crowding}
-> leads to stabilization of native structure due to the excluded volume effect ~\cite{cheung2005molecular}
-> can also destabilize proteins due to  ~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}. These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}. More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}. However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.

@@

\begin{table}[H]
\centering
\begin{tabular}{ l c c }
\hline
\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} \\
& Absolute & Cumulative \\ 
\hline
TufA & 19.7 & 19.7 \\
MetE & 11.6 & 31.4 \\
IcdA & 4.7  & 36.1 \\
AhpC & 4.1  & 40.2 \\
CspC & 4.0  & 44.2 \\
Ppa  & 2.9  & 47.0 \\
GapA & 2.1  & 49.1 \\
Eno  & 1.9  & 51.1 \\
\hline
\end{tabular}
\caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E. coli} K12 to which each protein that was selected to compose our cytoplasm model corresponds to.}
\label{tbl:protein_fraction}
\end{table}

@@

\subsubsection{Water fraction}
The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model. This number ranges from \colorbox{red!50}{XX to YY} in biological systems such as \textit{E. coli} cytoplasm. In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.

@@

We devised an iterative process that solves both problems simultaneously. We start with a box size $L$ that is too small to allow all components to fit in the box by random insertion. In our case, we started with $L = 30$ nm. Then, we allow 100 insertion trials for each element. If all trials fail for any of them, we start the whole process again with an empty cubic box that is larger by a step size $dL$ of 1 nm. We repeat this process until all insertions succeed. In our case, all insertions succeeded after increasing $L$ to 35 nm. Additionally, instead of adding only the protein, RNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded. Such droplets are taken from molecular dynamics simulations in which each component was previously equilibrated. The benefits of using such droplets are threefold: \textit{i.} it acts as natural protecting layer that prevents artificial contacts between the components that could be created due to the random insertions. \textit{ii.} it is a natural way to place water molecules and counter-ions in the simulation box around each component. \textit{iii.} the components are already pre-equilibrated, which will help us to perform the equilibration of the whole cytoplasm model.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking_volume.pdf} 
\includegraphics[scale=0.5]{box.pdf} 
% \includegraphics[scale=0.11]{box_big.png}
% \includegraphics[scale=0.11]{box_small_.png}
%\includegraphics[scale=0.5]{shrinking_density.pdf}

\caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$. The repeated experiments are shown in different colors. }
\end{figure}

@@

In order to investigate the effects of crowding in the \textit{E. coli} cytoplasm model on the structural integrity and dynamics of its elements, we constructed three completely independent cytoplasm models that have different orientations for each of its elements. We submitted these systems, each composed by more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}. We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element in order to represent the non-crowded, i.e. dilute, environment (please check the materials section for details about the parameters).

@@

Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment. The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Figure~\ref{fig:translational_diffusion}A). The ratio of these quantities is independent of the protein size and is always close to 0.13 (Figure~\ref{fig:translational_diffusion}B). However, the effect of crowding over the translation diffusion of tRNA is two times higher. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Figure~\ref{fig:translational_diffusion}B, crowder with 25 kDa).

\begin{figure}[H]
\includegraphics[scale=0.5]{msd.pdf}
\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
\caption{A)Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars). B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis.}
\label{fig:translational_diffusion}
\end{figure}

@@

\subsubsection{tRNA is aggregating}
After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Figure~\ref{fig:tRNA_aggregation}). Thus, in the next sections we will ignore the RNA molecules in the analyses about the structural integrity of the crowders. Additionally, in the last section we will further investigate such aggregation and we will show a way to prevent it.

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.}
\label{fig:tRNA_aggregation}
\end{figure}

@@

\subsubsection{Structural integrity of individual chains}
The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment. The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition. The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show a similar behavior, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition. Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues \colorbox{red!50}{DESCRIBE CONF. CHANGES}.
\begin{figure}[H]
\includegraphics[scale=0.5]{sasa.pdf}
\includegraphics[scale=0.5]{rmsd.pdf}

\caption{A) Chain interface area for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average bar value for single molecule simulations shows the sum over sasa of each protein chains averaged over the number of replica minus the sasa of the whole protein averaged over the number of replica. The average bar value for cytoplasm simulation shows the sum over sasa of each protein chains averaged over the number of replica and the abundance of the protein minus the sasa of the whole protein averaged over the number of replica and the abundance of the protein. The error bars show the standard errors. B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.}
\label{fig:structural_integrity_chain}
\end{figure}

% -------------------- STRUCTURAL INTEGRITY OF OLIGOMERS --------------------

\subsubsection{Structural integrity of oligomers}

% \colorbox{red!50}{WE STILL NEED TO MAKE THIS PLOT}

%\begin{figure}[H]
%\includegraphics[scale=0.5]{oligomer.pdf}
%\caption{Inter-chain hydrogen bonds (hb) for oligomers from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average hydrogen bonds for each oligomer in single molecule simulation is over number of chains and three replica. The average hydrogen bonds for each oligomer in cytoplasm simulation is over number of appearance of the chain in the oligomer in three replica. The error bars show the standard errors.}
%\end{figure}

@@

\subsubsection{Structural dynamics}
Proteins are slightly more flexible in the crowded environment. Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes, TufA ... as shown before (Figure~\ref{fig:rmsf}). This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected. In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition (Figure~\ref{fig:rmsf}). This is likely due to the nonspecific interactions and collisions that often happen under crowding conditions \textbf{\color{red}{REF}}.

@@

The initial detection of aggregation in our cytoplasm model was due to the out-lier behavior of the translational diffusion constant of tRNA when compared to the other crowders. However, tRNA is not causing such phenomenon. Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$. tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few "sources" of Mg$^{2+}$ in the whole system. We looked for ways of avoid aggregation with simulations of small systems composed by these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).

We found that completely protonating the phosphate groups of ATP and FBP is enough prevent their aggregation with Mg$^{2+}$ (Figure~\ref{fig:avoiding_aggregation}). While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$, this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Figure~\ref{fig:protonation}).

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_mg.pdf}
\caption{Radial distribution function (rdf) showing the probability of finding a metabolite or an RNA around MG2+. The inset plot shows the rdf for short distance ($ r < 0.7 \,\, nm$)}
\label{fig:avoiding_aggregation}
\end{figure}

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_mg_metabolites.pdf}
\caption{A) Radial distribution function (rdf) showing the probability of finding an ATP$^{-3}$ or an ATPH\textsubscript{3} around Mg$^{2+}$. B) Radial distribution function (rdf) showing the probability of finding an FBP$^{-4}$ or an FBPH\textsubscript{4} around Mg$^{2+}$. }
\label{fig:protonation}
\end{figure}

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an out-lier in our cytoplasm model. In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data. Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photo-bleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}. In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model. All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{yu2016biomolecular}. The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

All crowders are as compact in the cytoplasm model as they are in dilute conditions (Figure \ref{fig:structural_integrity_chain}A). Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Figure \ref{fig:structural_integrity_chain}B). Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Figure \ref{fig:rmsf}). This is in agreement with simulation~\cite{feig2011variable} and experimental data~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.

The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}

Magnesium cations were added to our cytoplasm model as counter-ions for tRNA and ATP. However, as the simulation progressed, FBP and ATP aggregated around Mg$^{2+}$ and, by consequence, around tRNA. After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models. We found that protonating the phosphate-containing metabolites is enough to reach this goal (Figure \ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results. The biggest impact of this approach is on long range interactions of these metabolites with other molecules due to the neutralization of their net charge. However, molecules in a crowded environment are restrained to smaller volumes and shorter inter molecular distances, which mitigates the effect of charge neutralization.

To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}. Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was "much slower" than expected. They attributed this observation to non-specific interactions between such metabolites and proteins. Maybe aggregation also contributed to their slower diffusion but it was not evident during \SI{140}{\nano\second}, which was the length of their longest simulation.

>>>>

\documentclass[journal=jacsat,manuscript=article]{achemso}
% \linespread{1.5}
\usepackage{amssymb,amsmath,graphics,epsfig}
% \usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

The intracellular microenvironment is markedly different from the \textit{in vitro} simplification from which most of our current knowledge about the biochemical and biophysical behavior of macromolecules was derived~\cite{Feig2017a}. Instead of simple and diluted, the \textit{in vivo} microenvironment is highly heterogeneous and concentrated, leading to spatial constrains that directly affects biological activities~\citep{ostrowska2019}. Macromolecular crowding directly influences the stability of macromolecules. There is an equilibrium between stabilization of their native structure due to the entropic component given by the excluded volume effect~\cite{cheung2005} and destabilizationg due to enthalpic component given by unspecific interactions with the crowders~\cite{Feig2011,miklos2011,Wang2012b}. Additionally, crowding affects the structure and dynamics of water around proteins, which has an indirect but important effect in their biological activities~\cite{Harada2012a,king2013}.

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluorescence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply Brownian Dynamics calculations to probe macromolecules immersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules~\cite{Ando2010}. These models usually use an implicit representation of the solvent~\cite{Mcguffee2010}. More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed of several copies of one or a few macromolecules~\cite{Wang2017c}. However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms~\cite{Feig2015}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations~\cite{Yu2016a}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.

@@

\begin{table}[H]
\centering
\begin{tabular}{ l c c c}
\hline
\multirow{2}{*}{Protein} & 	\multicolumn{2}{c}{Fraction [\%]} & \multirow{2}{*}{Oligomeric State}\\
& Absolute & Cumulative \\ 
\hline
TufA & 19.7 & 19.7 & 2-mer \\
MetE & 11.6 & 31.4 & 1-mer \\
IcdA & 4.7  & 36.1 & 2-mer \\
AhpC & 4.1  & 40.2 & 10-mer \\
CspC & 4.0  & 44.2 & 1-mer \\
Ppa  & 2.9  & 47.0 & 6-mer \\
GapA & 2.1  & 49.1 & 4-mer \\
Eno  & 1.9  & 51.1 & 2-mer \\
\hline
\end{tabular}
\caption{Absolute and cumulative fraction of the total abundance of non-ribosomal proteins in the cytosol of {\em E. coli} K12 to which each protein that was selected to compose our cytoplasm model corresponds to.}
\label{tbl:protein_fraction}
\end{table}

@@

\subsubsection{Water fraction}
The number of water molecules was calculated according to the desired biomolecular concentration, which is a parameter of the cytoplasm model. This number ranges from 300 to 400 $mg/ml$ in biological systems such as \textit{E. coli} cytoplasm~\cite{Zimmerman1991,Ellis2003a}. In our case, we choose the biomolecular concentration of 30\%, that is, the number of molecules necessary to reach a ratio of total biomolecular mass to water mass of 30\% was inserted into the cytoplasm model.

@@

We devised an iterative process that solves both problems simultaneously. We start with a box size $L$ that is too small to allow all components to fit in the box by random insertion. In our case, we started with $L = 30$ nm. Then, we allow 100 insertion trials for each element. If all trials fail for any of them, we start the whole process again with an empty cubic box that is larger by a step size $dL$ of 1 nm. We repeat this process until all insertions succeed. In our case, all insertions succeeded after increasing $L$ to 35 nm. Additionally, instead of adding only the protein, tRNA or metabolite molecules in the empty box in each trial, we actually add a droplet of water and counter ions in which the molecule of interest is embedded. Such droplets are taken from molecular dynamics simulations in which each component was previously equilibrated. The benefits of using such droplets are threefold: \textit{i.} it acts as natural protecting layer that prevents artificial contacts between the components that could be created due to the random insertions. \textit{ii.} it is a natural way to place water molecules and counter-ions in the simulation box around each component. \textit{iii.} the components are already pre-equilibrated, which will help us to perform the equilibration of the whole cytoplasm model.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking.pdf} 
\caption{A) The volume of the box with length size of $L=35 \,\, nm$ reduces during the energy minimization in the first $300 \,\, ps$ B) The density of the same box increases within the first $300\,\, ps$. The repeated experiments are shown in different colors. }
\end{figure}

@@

In order to investigate the effects of crowding in the \textit{E. coli} cytoplasm model on the structural integrity and dynamics of its elements, we constructed three completely independent cytoplasm models that have different orientations for each of its elements. We submitted these systems, each composed of more than 1.5 million atoms, to molecular dynamics simulations of \SI{1}{\micro\second}. We also performed \SI{200}{\nano\second} molecular dynamics simulations for each isolated element (proteins and tRNA) in order to represent the non-crowded, i.e. dilute environment (please check the materials section for details about the parameters).

@@

Under crowded conditions, molecules are constrained to a smaller volume due to collisions with the other components of the crowded environment. The extent of this effect is evident when we compare the translational diffusion constant of each component in the cytoplasm model, $D_{cyto}$, and in a dilute system, $D_{dil}$ (Fig.~\ref{fig:translational_diffusion}A). The ratio of these quantities is independent of the protein size and is always close to 0.13 (Fig.~\ref{fig:translational_diffusion}B). However, the effect of crowding on the translational diffusion of tRNA is two times larger. Its $D_{cyto}/D_{dil}$ is 0.07, indicating that its movement is specially constrained in the crowded environment (Fig.~\ref{fig:translational_diffusion}B, crowder with 25 kDa).

\begin{figure}[H]
\includegraphics[scale=0.5]{msd.pdf}
\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
\caption{A) Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars). B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis.}
\label{fig:translational_diffusion}
\end{figure}

@@

\subsubsection{tRNA is aggregating}
After inspecting the trajectories of the cytoplasm models, we found that the reason why the translational diffusion constant for tRNA was reduced to a greater extent than for the other components of the cytoplasm model is that it is forming aggregates with Mg$^{2+}$, ATP and FBP (Fig.~\ref{fig:tRNA_aggregation}). Thus, in the next sections we will ignore the tRNA molecules in the analyses about the structural integrity of the crowders. Additionally, in the last section we will further investigate such aggregation and we will show a way to prevent it.

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
\caption{Radial distribution function showing the probability of finding metabolites or Magnesium cations around tRNA molecules in the cytoplasm model.}
\label{fig:tRNA_aggregation}
\end{figure}

@@

\subsubsection{Structural integrity}
The structure of the individual chains of the crowders that were used in our cytoplasm model were not affected to a great extent by the crowded environment. %\st{The solvent accessible surface area (SASA) value for all crowders in the cytoplasm model is within 5\% of its value under the dilute condition.} 
The root mean square deviation (RMSD) considering only their C$\upalpha$ atoms show that crowders remain intacts under crowded condition, except for TufA, MetE and IcdA, which have RMSD values 105, 25 and 33\% higher in the crowded environment than in the dilute condition. Visual inspection of their trajectories showed that there is partial unfolding of their N- or C-terminal residues. 
\begin{figure}[H]
\includegraphics[scale=0.5]{rmsd.pdf}
\includegraphics[scale=0.5]{sasa.pdf}

\caption{A) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors. B) Chain interface area for oligomers from single molecule simulations (blue bars) and cytoplasm simulations (orange bars). The average value from single molecule simulation is over three replica. The average value from cytoplasm simulation is over the number of proteins and three replica. The error bars show the standard errors. }
\label{fig:structural_integrity_chain}
\end{figure}

@@

\subsubsection{Structural dynamics}
Proteins are slightly more flexible in the crowded environment. Root mean square fluctuation (RMSF) calculations show that the flexibility profile of the proteins is not affected by crowding, except in the cases in which there are conformational changes (TufA, MetE and IcdA) (Fig.~\ref{fig:rmsf}). This is in agreement with our results showing that the structural integrity of the proteins is not significantly affected (Fig.~\ref{fig:structural_integrity_chain}A). 
%In addition to that, our results show that the proteins are slightly more flexible in the crowded environment than in the dilute condition. 
This is likely due to the nonspecific interactions that often happen under crowding conditions~\cite{Feig2011, miklos2011, Wang2012b}. The effec of crowded environment on inter-chain interaction of type hydrophobic, hydrogen bonding and salt bridges can be assessed by integrity of oligomeric chains. This is measured as chain interface area (Fig.~\ref{fig:structural_integrity_chain}B). Slightly lower buried area in cytoplasm simulations manifests destabilizing effect of crowded environment on chains of oligomers except except IcdA, GapA and Eno (Fig.~\ref{fig:structural_integrity_chain}B). These three oligomers also show the least changes in their RMSF profiles (Fig.~\ref{fig:rmsf}).

@@

The initial detection of aggregation in our cytoplasm model was due to the outlier behavior of the translational diffusion constant of tRNA when compared to the other crowders. However, tRNA is not causing such phenomenon (Fig.~\ref{fig:avoiding_aggregation}A). Instead, it is triggered by the exaggerated interaction between the phosphate-containing metabolites, ATP and FBP, with Mg$^{2+}$. tRNA gets involved in the aggregates because Mg$^{2+}$ is present in the box only as counter-ion for ATP and tRNA, and so tRNA is one of the few ``sources'' of Mg$^{2+}$ in the whole system. We looked for ways of avoiding aggregation by simulations of small systems composed of these metabolites in high concentration, Mg$^{2+}$ and water (please check the methods section for details about the parameters we used).

% We found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C). While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), this is not observed for ATPH\textsubscript{3} and FBPH\textsubscript{4}(Fig.~\ref{fig:avoiding_aggregation}B-C).

While RDFs show that ATP$^{-3}$ and FBP$^{-4}$ aggregate around Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}A), we found that completely protonating the phosphate groups of ATP and FBP is enough to prevent their aggregation with Mg$^{2+}$ (Fig.~\ref{fig:avoiding_aggregation}B-C).

\begin{figure}[H]
\hspace*{-2cm}\includegraphics[scale=0.5]{rdf_mg.pdf}
\caption{Radial distribution function (rdf) showing the probability of finding A) metabolites, B) ATP$^{-3}$ or an ATPH\textsubscript{3}, C) FBP$^{-4}$ or an FBPH\textsubscript{4} around MG2+. The averages are over three replica and the shaded regions are the standard errors.}
\label{fig:avoiding_aggregation}
\end{figure}

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an outlier in our cytoplasm model. In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data. Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photo-bleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions~\cite{andrews2013}. Moreover, it has been reported that simulations of highly concentrated environment leads to aggregation by using additive force fields~\cite{Petrov2014a,Abriata2015a, Nawrocki2017a}. Several solutions has been proposed to eliminate aggregation~\cite{Best2014a,Piana2015a, Bashardanesh2018b}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields~\cite{Yu2016a}, whose solvent-solute interactions are strengthened according to the solution proposed by Best {\em et al.}~\cite{Best2014a}. In this work we use the AMBER force fields~\cite{Best2014a} to study crowding effects on a cytoplasm model. All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed of 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively~\cite{Yu2016a}. The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

All crowders are as compact in the cytoplasm model as they are in dilute conditions (Fig.~\ref{fig:structural_integrity_chain}). Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Fig.~\ref{fig:structural_integrity_chain}). Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Fig.~\ref{fig:rmsf}). This is in agreement with simulation~\cite{Feig2011} and experiment~\cite{miklos2011,Wang2012b} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.

Having variation of oligomeric state of the crowders in our model (Table~\ref{tbl:protein_fraction}) enabled us to study the effec of crowding on inter-chain interactions. We found out that the oligomers are either slightly destabilized, or no effect is seen (GapA and Eno) (Fig.~\ref{fig:structural_integrity_chain}B). This is in agreement with rmsf results (Fig.~\ref{fig:rmsf}).

Magnesium cations were added to our cytoplasm model as counter-ions for tRNA and ATP. However, as the simulation progressed, FBP and ATP aggregated around Mg$^{2+}$ and, by consequence, around tRNA. After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models. We found that protonating the phosphate-containing metabolites is enough to reach this goal (Fig.~\ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results. The biggest impact of this approach is on long range interactions of these metabolites with other molecules due to the neutralization of their net charge. However, molecules in a crowded environment are restrained to smaller volumes and shorter inter molecular distances, which mitigates the effect of charge neutralization.

To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics~\cite{Yu2016a}. Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was ``much slower'' than expected. They attributed this observation to non-specific interactions between such metabolites and proteins. Maybe aggregation also contributed to their slower diffusion but it was not evident during \SI{140}{\nano\second}, which was the length of their longest simulation.
